---
title: "DiffBind_CUTnRUN_H3K4me3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DiffBind)
setwd("/mnt/dolab/group_folders/Personal/Christina/ernst01/spermatogenesis/cut_and_run/DiffBind")
# Read in Sample Sheet
H3K4me3_all <- dba(sampleSheet = "DiffBind_CUTnRUN_H3K4me3_all.csv")

H3K4me3_all

```


```{r}
# Occupancy based heatmap, depending on peaks called in different datasets
plot(H3K4me3_all)
```


```{r}
# Take read counts into consideration using dba.count using the raw reads
H3K4me3_counts <- dba.count(H3K4me3_all, score = DBA_SCORE_READS, bParallel = TRUE)

plot(H3K4me3_counts)
```
P18 samples and P20 spermatid sample don't cluster well with other samples. These three samples had a large number of called peaks and very low FRiP (<0.2). Probably best to remove these samples. 

```{r}
# Read in samples from P20 to P30

H3K4me3 <- dba(sampleSheet = "DiffBind_CUTnRUN_H3K4me3_P20-P30.csv")
H3K4me3
plot(H3K4me3)

# Consensus peakset for different cell types - present in at least half of samples of the same cell type
Consensus <- dba.peakset(H3K4me3, consensus = DBA_TISSUE, minOverlap = 0.5)
Consensus <- dba(Consensus, mask=Consensus$masks$Consensus, minOverlap=1)
# Plot overlap between these peaksets
dba.plotVenn(Consensus, Consensus$masks$Consensus)

# Extract consensus peaks
Consensus_peaks <- dba.peakset(Consensus, bRetrieve = TRUE)


# Count reads only across consensus peakset
Consensus_counts <- dba.count(H3K4me3, peaks=Consensus_peaks, score = DBA_SCORE_TMM_READS_FULL)
plot(Consensus_counts)
```

For differential binding between cell types, extract individual timepoints from dba object and compare between spermatocytes and spermatids from the same animal. Possibly also compare spermatocytes and spermatids separately between timepoints. 

```{r}
# Extract P22 samples
P22 <- dba(H3K4me3, mask=H3K4me3$masks$P22)
# Only keep peaks that are present in both replicates for each cell type
P22_consensus <- dba.peakset(P22, consensus = DBA_TISSUE, minOverlap = 2)
P22_consensus <- dba(P22_consensus, P22_consensus$masks$Consensus, minOverlap = 1)

# Plot overlap between these peaksets
dba.plotVenn(P22_consensus, P22_consensus$masks$Consensus)

# Extract consensus peaks
P22_consensus_peaks <- dba.peakset(P22_consensus, bRetrieve = TRUE)

# Counts over consensus peaks for P22
P22_consensus_counts <- dba.count(P22, peaks = P22_consensus_peaks, score = DBA_SCORE_RPKM)

# Establish contrast between samples for differential binding analysis between cell types
P22_contrast <- dba.contrast(P22_consensus_counts, categories = DBA_TISSUE, minMember =2)
P22_dba <- dba.analyze(P22_contrast, method = DBA_DESEQ2)

#Extract all reagions
P22_db = dba.report(P22_dba, th=1)
P22_DB = as.data.frame(P22_db)

library(ggplot2)
library(dplyr)

P22_ggplot <- P22_DB
# Minimum fold change of 2.5
P22_ggplot <- mutate(P22_ggplot, Fold_change = Fold < -2.5 | Fold >2.5)
# FDR <0.1
P22_ggplot <- mutate(P22_ggplot, Significant = FDR <0.1 & Fold_change == "TRUE")

#Pearson correlation
cor.test(x = P22_ggplot[,8], y=P22_ggplot[,7], method = "pearson")
lm_H3K4me3 <- lm(P22_ggplot[,8] ~ P22_ggplot[,7])

ggplot(P22_ggplot, aes(x = Conc_spermatocytes, y = Conc_spermatids, colour= Significant)) + geom_point(size = 1.5, shape = 19) + ggtitle("H3K4me3 - Spermatocytes vs. Spermatids - P22") + labs(x = "Spermatocytes", y="Spermatids") + scale_colour_manual(values = c("gray28", "darkorchid4"))
```
