---
title: "Hirota data"
author: "Nils Eling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, 
        encoding = encoding, output_file = '/Users/nils/Dropbox (Cambridge University)/SST_spermatocytes/Revisions/Results/Chip-Seq/Hirota.html') })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script analyses K9 signal in promoters of genes that are differentially regulated after Setdb1 KO.
It also aims at finding motivs for Setdb1 binding.

# Load data and libraries

```{r data, message=FALSE}
library(rtracklayer)
library(EnrichedHeatmap)
library(GenomicFeatures)
library(EnsDb.Mmusculus.v79)
library(TxDb.Mmusculus.UCSC.mm10.ensGene)
library(plyr)
library(reshape2)
library(pheatmap)
library(viridis)
library(plot3D)
library(ggsci)
library(cowplot)
library(openxlsx)
library(csaw)
library(edgeR)
library(Rsamtools)
library(Gviz)
source("../../Functions/auxiliary.R")

# Read in gene annotations
genenames <- read.table("../../Data/Mouse_genes.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
rownames(genenames) <- genenames$Gene.stable.ID

# Generate feature annotation
prom <- promoters(genes(EnsDb.Mmusculus.v79))
gene.body <- genes(EnsDb.Mmusculus.v79)
prom <- prom[seqnames(prom) %in% c(as.character(1:19), "X", "Y", "MT"),]
seqlevels(prom) <- c(as.character(1:19), "X", "Y", "MT")
gene.body <- gene.body[seqnames(gene.body) %in% c(as.character(1:19), "X", "Y", "MT"),]
seqlevels(gene.body) <- c(as.character(1:19), "X", "Y", "MT")

# K9 files P26
bam.files.K9.P26 <- list.files("../../../../Dropbox (Cambridge University)/SST_spermatocytes/Analysis/CnR/bam/H3K9me3/", full.names = TRUE,
                        pattern = paste(".+JP26.+bam$", sep = ""))

# Blacklisted regions
black <- import("http://mitra.stanford.edu/kundaje/akundaje/release/blacklists/mm10-mouse/mm10.blacklist.bed.gz")
seqlevels(black) <- sub("chr", "", seqlevels(black))
seqnames(black) <- sub("chr", "", seqnames(black))

# Parameters for reading bam files
param <- readParam(dedup=TRUE, minq=10, discard=black, pe="both", max.frag=1000)

# Bin the genome in 1000Kb windows
bins.K9.P26 <- windowCounts(bam.files.K9.P26, bin = TRUE, width = 1000, param=param)

# Read in genes that are differnetially expressed between WT and Setdb1 KO
Up.genes <- read.xlsx(xlsxFile = "https://ars.els-cdn.com/content/image/1-s2.0-S1534580718308256-mmc2.xlsx",
                      sheet = 1)
colnames(Up.genes) <- Up.genes[2,]
Up.genes <- Up.genes[-c(1,2),]

Down.genes <- read.xlsx(xlsxFile = "https://ars.els-cdn.com/content/image/1-s2.0-S1534580718308256-mmc2.xlsx",
                      sheet = 2)
colnames(Down.genes) <- Down.genes[2,]
Down.genes <- Down.genes[-c(1,2),]
```

# Count reads in promoter and gene body regions

```{r}
prom.K9 <- regionCounts(bam.files.K9.P26, regions = prom, param=param)
gene.K9 <- regionCounts(bam.files.K9.P26, regions = gene.body, param=param)

# Generate RPM per promoter - each promoter is 2200bp wide
cur_counts.prom.K9 <- assays(prom.K9)$counts
cur_rpm.prom.K9 <- t(t(cur_counts.prom.K9)/(colSums(assays(bins.K9.P26)$counts)/1000000))

# FPKM normalized counts per gene
cur_counts.gene.K9 <- assays(gene.K9)$counts
cur_rpm.gene.K9 <- t(t(cur_counts.gene.K9)/(colSums(assays(bins.K9.P26)$counts)/1000000))
cur_rpkm.gene.K9 <- cur_rpm.gene.K9/(gene.body@ranges@width/1000)

# Compute average between spermatocytes and spermatids
df <- data.frame(Spermatocytes.prom.K9 = rowMeans(cur_rpm.prom.K9[,c(1,3)]),
                 Spermatids.prom.K9 = rowMeans(cur_rpm.prom.K9[,c(2,4)]),
                 Spermatocytes.gene.K9 = rowMeans(cur_rpkm.gene.K9[,c(1,3)]),
                 Spermatids.gene.K9 = rowMeans(cur_rpkm.gene.K9[,c(2,4)]))
df$Symbol <- genenames[rownames(df),2]
df$Up_genes <- df$Symbol %in% Up.genes$`gene name`
df$Down_genes <- df$Symbol %in% Down.genes$`gene name`
df$chr <- as.character(seqnames(prom))
```

Visualize H3K9me3 signal of genes that are up-regulated after Setdb1 KO in spermatids and spermatocytes

```{r}
# Spermatocytes
# Across all promoters
ggplot(df) + geom_boxplot(aes(Up_genes, log(Spermatocytes.prom.K9 + 1), fill = Up_genes))

# Across all gene bodies
ggplot(df) + geom_boxplot(aes(Up_genes, log(Spermatocytes.gene.K9 + 1), fill = Up_genes))

# Promoters on X chromosome
ggplot(df[df$chr == "X",]) + geom_boxplot(aes(Up_genes, log(Spermatocytes.prom.K9 + 1), 
                                              fill = Up_genes))

# Genes on X chromosome
ggplot(df[df$chr == "X",]) + geom_boxplot(aes(Up_genes, log(Spermatocytes.gene.K9 + 1), 
                                              fill = Up_genes))

# Spermatids
# Across all promoters
ggplot(df) + geom_boxplot(aes(Up_genes, log(Spermatids.prom.K9 + 1), fill = Up_genes))

# Across all gene bodies
ggplot(df) + geom_boxplot(aes(Up_genes, log(Spermatids.gene.K9 + 1), fill = Up_genes))

# Promoters on X chromosome
ggplot(df[df$chr == "X",]) + geom_boxplot(aes(Up_genes, log(Spermatids.prom.K9 + 1), 
                                              fill = Up_genes))

# Genes on X chromosome
ggplot(df[df$chr == "X",]) + geom_boxplot(aes(Up_genes, log(Spermatids.gene.K9 + 1), 
                                              fill = Up_genes))
```

Visualize H3K9me3 signal of genes that are down-regulated after Setdb1 KO in spermatids and spermatocytes

```{r}
# Spermatocytes
# Across all promoters
ggplot(df) + geom_boxplot(aes(Down_genes, log(Spermatocytes.prom.K9 + 1), fill = Down_genes))

# Across all gene bodies
ggplot(df) + geom_boxplot(aes(Down_genes, log(Spermatocytes.gene.K9 + 1), fill = Down_genes))

# Promoters on X chromosome
ggplot(df[df$chr == "X",]) + geom_boxplot(aes(Down_genes, log(Spermatocytes.prom.K9 + 1), 
                                              fill = Down_genes))

# Genes on X chromosome
ggplot(df[df$chr == "X",]) + geom_boxplot(aes(Down_genes, log(Spermatocytes.gene.K9 + 1), 
                                              fill = Down_genes))

# Spermatids
# Across all promoters
ggplot(df) + geom_boxplot(aes(Down_genes, log(Spermatids.prom.K9 + 1), fill = Down_genes))

# Across all gene bodies
ggplot(df) + geom_boxplot(aes(Down_genes, log(Spermatids.gene.K9 + 1), fill = Down_genes))

# Promoters on X chromosome
ggplot(df[df$chr == "X",]) + geom_boxplot(aes(Down_genes, log(Spermatids.prom.K9 + 1), 
                                              fill = Down_genes))

# Genes on X chromosome
ggplot(df[df$chr == "X",]) + geom_boxplot(aes(Down_genes, log(Spermatids.gene.K9 + 1), 
                                              fill = Down_genes))
```

Find genes that have high signal for K9 in spermatocytes and are up-regulated after Setdb1 KO.

```{r}

```