---
title: "Variability over pseudotime"
author: "Nils Eling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, 
        encoding = encoding, output_file = '/Users/nils/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/Variability/VariabilityOverPT.html') })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I ran the BASiCS model on each cell population individually using the replicates to estimate technical variance. For each gene I fitted a linear regression between each iteration against the group number as response variable. This gives us a post-hoc posterior distribution of the slope of the linear model.

# Variability over pseudotime

```{r}
# Load libraries and data
library(BASiCS)
library(coda)
library(pheatmap)
genenames <- read.table("/Users/nils/Google Drive File Stream/My Drive/Christina/Tc1_single_cell/data/raw_reads/10X/do15983/Mus_human_chr21/genes.tsv", sep = "\t", stringsAsFactors = FALSE)
rownames(genenames) <- genenames[,1]
fits <- readRDS("/Users/nils/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/10X/Variability/data/fits_B6.rds")
fits <- fits[,!is.na(genenames[colnames(fits), 2])]

# Visualize some trace plots for the MCMCs
plot(mcmc(fits[,1]))
plot(mcmc(fits[,10]))
plot(mcmc(fits[,100]))
plot(mcmc(fits[,1000]))
```

The trace plots show convergence for the slope - we can use this for posterior tail propability testing.

```{r}
# Estimate the EFDR
TestSlope <- BASiCS:::HiddenThresholdSearchTestDE(ChainLFC = fits, Epsilon = 0,
                                                  ProbThreshold = NULL, GenesSelect = NULL,
                                                  EFDR = 0.1, Task = "Slope testing")

OptThreshold <- TestSlope$OptThreshold[1]

slope <- colMedians(fits)

# Build output dataframe
df.out <- data.frame(row.names = colnames(fits),
                     Genenames = genenames[colnames(fits),2],
                     Slope = slope,
                     Probability = TestSlope$Prob,
                     Result = ifelse(TestSlope$Prob > OptThreshold & slope > 0, "UpWithPT",
                                     ifelse(TestSlope$Prob > OptThreshold & slope < 0, "DownWithPT", "NoDiff")))
```

Now we can order the genes based on slope for increasing and decreasing variability over pseudotime.

```{r}
# Genes that increase variability with pseudotime
head(df.out[order(df.out$Slope, decreasing = TRUE),], n = 100)
# Genes that decrease variability with pseudotime
head(df.out[order(df.out$Slope, decreasing = FALSE),], n = 100)

write.csv(df.out[order(df.out$Slope, decreasing = TRUE),], "Results/VariabilityOverPT.csv")
```

It's interesting to plot the slope versus mean expression.

```{r}
norm <- readRDS("../data/10X_data/Tc0_young_norm.rds")
clusters <- read.csv("Results/Clusters_Tc0_young.csv", stringsAsFactors = FALSE)

norm <- norm[genenames[colnames(fits), 2],
             clusters$OrderByPT != 1 & clusters$OrderByPT != 2]

plot(log10(rowMeans(as.matrix(norm))), df.out$Slope, pch = 16,
     xlab = "Mean expression", ylab = "Slope",
     col = ifelse(df.out$Result == "UpWithPT", "dark red",
                  ifelse(df.out$Result == "DownWithPT", "dark blue",
                  "grey")))

df.out[log10(rowMeans(as.matrix(norm))) < 1 &
         df.out$Slope > 2.5,]
```

# Variability on X chromosome

```{r}
chrX <- read.table("../data/chrX.txt", sep = "\t", stringsAsFactors = FALSE, header = TRUE)
df.out[rownames(df.out) %in% chrX$Gene.stable.ID,]
```