---
title: "Variability over pseudotime"
author: "eling01 Eling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, 
        encoding = encoding, output_file = '/Users/eling01/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/Variability/VariabilityOverPT.html') })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I ran the BASiCS model on each cell population individually using the replicates to estimate technical variance. For each gene I fitted a linear regression between each iteration against the group number as response variable. This gives us a post-hoc posterior distribution of the slope of the linear model.

# Variability over pseudotime

```{r, message=FALSE}
# Load libraries and data
library(BASiCS)
library(coda)
library(pheatmap)
library(openxlsx)
library(cowplot)
library(ggplot2)
library(rtracklayer)
library(EnrichedHeatmap)
library(GenomicFeatures)
library(EnsDb.Mmusculus.v79)
library(scater)
library(plyr)
library(reshape2)
library(pheatmap)
library(viridis)
source("../../Functions/auxiliary.R")

# Read in sce
sce <- readRDS("/Users/eling01/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/data/10X_data/SCE_all_clusters.rds")
sce <- sce[,grepl("B6", colData(sce)$Sample) &
             colData(sce)$AnnotatedClusters %in% 
             c("S1", "S2", "S3", "S4", "S5", "S6", "S7", "S8", "S9")]
sce <- normalize(sce)

# Generate feature annotation
prom <- promoters(genes(EnsDb.Mmusculus.v79),
                  upstream = 1500, downstream = 500)
prom <- prom[seqnames(prom) %in% c(as.character(1:19), "X", "Y", "MT"),]
tss <- promoters(genes(EnsDb.Mmusculus.v79), 
                 upstream = 0, downstream = 1)
tss <- tss[seqnames(tss) %in% c(as.character(1:19), "X", "Y", "MT"),]
seqlevels(tss) <- c(as.character(1:19), "X", "Y", "MT")

# Read in blacklist file
blacklist <- import("/Users/eling01/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/Cnr/Blacklist_mm10.bed")
seqlevels(blacklist) <- sub("^chr", '', seqlevels(blacklist))

# Read in gene names
genenames <- read.table("/Users/eling01/Google Drive File Stream/My Drive/Christina/Tc1_single_cell/data/raw_reads/10X/do15983/Mus_human_chr21/genes.tsv", sep = "\t", stringsAsFactors = FALSE)
rownames(genenames) <- genenames[,1]

# Read in post-hoc distribution of slopes based on epsilons
fits.epsilon <- readRDS("/Users/eling01/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/10X/Variability/data/fits_epsilon_B6.rds")
fits.mu <- readRDS("/Users/eling01/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/10X/Variability/data/fits_mu_B6.rds")

# Read in medians of epsilons and mus
mus <- readRDS("/Users/eling01/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/10X/Variability/data/mus_B6.rds")
epsilons <- readRDS("/Users/eling01/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/10X/Variability/data/epsilons_B6.rds")

# Remove coloumns containins NAs
select <- !is.na(colSums(fits.epsilon)) & !is.na(colSums(fits.mu))
fits.epsilon <- fits.epsilon[,select]
fits.mu <- fits.mu[,select]

# Visualize some trace plots for the MCMCs
plot(mcmc(fits.epsilon[,1]))
plot(mcmc(fits.epsilon[,10]))
plot(mcmc(fits.epsilon[,100]))
plot(mcmc(fits.epsilon[,1000]))
```

The trace plots show convergence for the slope - we can use this for posterior tail propability testing.

## Testing increases of decreases over PT

### Testing for fits on epsilons

```{r}
# Estimate the EFDR
TestSlope.eps <- BASiCS:::HiddenThresholdSearchTestDE(ChainLFC = fits.epsilon, 
                                        Epsilon = 0,
                                        ProbThreshold = NULL, GenesSelect = NULL,
                                        EFDR = 0.1, Task = "Slope testing")

OptThreshold.eps <- TestSlope.eps$OptThreshold[1]

slope.eps <- colMedians(fits.epsilon)
```

### Testing for fits on mus

```{r}
# Estimate the EFDR
TestSlope.mu <- BASiCS:::HiddenThresholdSearchTestDE(ChainLFC = fits.mu, 
                                        Epsilon = 0,
                                        ProbThreshold = 0.8, GenesSelect = NULL,
                                        EFDR = 0.1, Task = "Slope testing")

OptThreshold.mus <- TestSlope.mu$OptThreshold[1]

slope.mu <- colMedians(fits.mu)

# Build output dataframe
df.out <- data.frame(row.names = colnames(fits.mu),
                     Genenames = genenames[colnames(fits.mu),2],
                     Slope.mu = slope.mu,
                     Probability.mu = TestSlope.mu$Prob,
                     Result.mu = ifelse(TestSlope.mu$Prob > OptThreshold.mus & 
                                       slope.mu > 0, "MeanUp",
                                     ifelse(TestSlope.mu$Prob > OptThreshold.mus & 
                                              slope.mu < 0, "MeanDown", "NoDiff")),
                     Slope.epsilon = slope.eps,
                     Probability.epsilon = TestSlope.eps$Prob,
                     Result.epsilon = ifelse(TestSlope.eps$Prob > OptThreshold.eps & 
                                       slope.eps > 0, "VariabilityUp",
                                     ifelse(TestSlope.eps$Prob > OptThreshold.eps & 
                                              slope.eps < 0, "VariabilityDown", "NoDiff")))

# Order genes based on their increase in variability over PT
df.out <- df.out[order(df.out$Slope.epsilon, decreasing = TRUE),]
df.out$regulation <- paste(df.out$Result.mu, 
                           df.out$Result.epsilon, sep = "_and_")

write.xlsx(df.out, "/Users/eling01/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/10X/Results/VariabilityOverPT.xlsx")
```

### Classification of genes into 8 groups

- Increase in variability, no change in mean expression
- Decrease in variability, no change in mean expression
- Increase in variability, increase in mean expression
- Decrease in variability, increase in mean expression
- Increase in variability, decrease in mean expression
- Decrease in variability, decrease in mean expression
- No change in variability, increase in mean expression
- No change in variability, decrease in mean expression
- No change in variability, no change in mean expression


```{r}
ggplot(df.out) + geom_point(aes(Slope.mu, Slope.epsilon, colour = regulation)) +
  scale_color_manual(values = c("#fddbc7", "#d7301f", "#7f0000", "#d1e5f0", "#0570b0", "#081d58", "#f7fcfd", "#f4a582", "#92c5de")) + xlim(c(-20, 20)) +
  xlab("Slope, Mean expression vs PT") + ylab("Slope, Variability vs PT")
```

### Order cells along the principle curve

```{r}
# Compute HVG
HVgenes <- HVG(sce)

# Compute PCA
pca <- prcomp(t(logcounts(sce)[HVgenes,]))

# Pseudo rank
prank <- PT(rd = pca$x[,1:2], clusters = colData(sce)$AnnotatedClusters,
            col_vector = metadata(sce)$color_vector)

# DPT 
#dpt <- diffusionPT(sce, HVgenes, clusters = colData(sce)$AnnotatedClusters,
#            col_vector = metadata(sce)$color_vector)
```

### Visualize example genes per category

```{r}
# Increase in mean expression, no change in variability
head(df.out[order(df.out$Slope.mu, decreasing = TRUE),], 50)

# Gpx4
cur.gene <- "Prm1"
cur.Z <- (logcounts(sce)[rowData(sce)$Symbol == cur.gene,] -
             mean(logcounts(sce)[rowData(sce)$Symbol == cur.gene,]))/
  sd(logcounts(sce)[rowData(sce)$Symbol == cur.gene,])

cur_p <- ggplot(data.frame(PT = prank[,"lambda"],
                  gene = cur.Z,
                  cell_type = colData(sce)$AnnotatedClusters)) + 
  geom_point(aes(PT, gene, colour = cell_type)) + 
  xlab("Pseudotime") + ylab(paste(cur.gene, ", z-score", sep = "")) +
  geom_abline(slope = 0, intercept = 0, colour = "dark red") + 
  scale_color_manual(values = metadata(sce)$color_vector) + 
  guides(color = FALSE)

cur_p.box <- ggplot(data.frame(PT = prank[,"lambda"],
                  gene = cur.Z,
                  cell_type = colData(sce)$AnnotatedClusters)) + 
  geom_boxplot(aes(cell_type, gene, colour = cell_type)) + 
  xlab("Pseudotime") + ylab(paste(cur.gene, ", z-score", sep = "")) +
  geom_abline(slope = 0, intercept = 0, colour 
              
              
              = "dark red") + 
  scale_color_manual(values = metadata(sce)$color_vector) + 
  guides(color = FALSE)

cur_p.var <- ggplot(data.frame(gene = epsilons[rowData(sce)$ID[rowData(sce)$Symbol == cur.gene],
                                   c("S1", "S2", "S3", "S4", "S5", "S6", "S7", "S8", "S9")],
                  cell_type = c("S1", "S2", "S3", "S4", "S5", "S6", "S7", "S8", "S9"))) + 
  geom_point(aes(cell_type, gene, colour = cell_type), size = 5) + 
  xlab("Pseudotime") + ylab(paste(cur.gene, ", Variability", sep = "")) +
  geom_abline(slope = 0, intercept = 0, colour = "dark red") + 
  scale_color_manual(values = metadata(sce)$color_vector) + 
  guides(color = FALSE)

```

# Rank genes by mean variability

```{r ranking}
# Rank the genes based on the epsilons from S1-S14
cur_eps <- epsilons[,paste("S", 1:14, sep = "")]
cur_eps <- cur_eps[!(sum(t(apply(cur_eps, 1, is.na))) == ncol(cur_eps)),]

cur_eps.ranked <- cur_eps[order(rowMeans(cur_eps, na.rm = TRUE), decreasing = TRUE),]
```

## Visualize 100 sample genes

```{r sample}
set.seed(123456)
sam <- sample(1:nrow(cur_eps.ranked), 100)
sam <- sam[order(sam)]

pheatmap(cur_eps.ranked[sam,], cluster_rows = FALSE, cluster_cols = FALSE, 
         color = colorRampPalette(c("dark blue", "white", "dark red"))(100), 
         labels_row = rowData(sce)$Symbol[match(rownames(cur_eps.ranked)[sam], rowData(sce)$ID)],
         breaks = seq(-7,7, length.out = 100))
```

## Visualize peaks for these genes

```{r peaks}
cur_files <- list.files("/Users/eling01/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/CnR/bedgraphs/H3K4me3/", 
                        full.names = TRUE)

# H3K4me3 spermatids day 24
cur_file <- import(cur_files[12])

# Exclude blacklisted regions
cur_removed <- overlapsAny(cur_file, blacklist)
cur_file <- cur_file[!cur_removed]
cur_file <- cur_file[seqnames(cur_file) %in%
                         c(as.character(1:19), "X", "Y", "MT"),]
seqlevels(cur_file) <- c(as.character(1:19), "X", "Y", "MT")


cur_genes <- rownames(cur_eps.ranked)[sam]
cur_tss <- tss[match(cur_genes, tss@elementMetadata$gene_id)]

cur_mat = normalizeToMatrix(cur_file, cur_tss, 
                        value_column = "score",
                        extend = 5000, mean_mode = "w0", 
                        w = 50, keep = c(0,0.99))

EnrichedHeatmap(cur_mat, col = c("white", "red", "black"), row_order = 1:length(cur_genes))

# Visualize top 100 and bottom 100 genes
cur_genes <- rownames(cur_eps.ranked)[1:100]
cur_tss <- tss[match(cur_genes, tss@elementMetadata$gene_id)]

cur_mat.highVar = normalizeToMatrix(cur_file, cur_tss, 
                        value_column = "score",
                        extend = 5000, mean_mode = "w0", 
                        w = 50, keep = c(0,0.99))

EnrichedHeatmap(cur_mat.highVar, col = c("white", "red", "black"))

cur_genes <- tail(rownames(cur_eps.ranked), 100)
cur_tss <- tss[tss@elementMetadata$gene_id %in% cur_genes]

cur_mat.lowVar = normalizeToMatrix(cur_file, cur_tss, 
                        value_column = "score",
                        extend = 5000, mean_mode = "w0", 
                        w = 50, keep = c(0,0.99))

EnrichedHeatmap(cur_mat.lowVar, col = c("white", "red", "black"))


```

# Plot mean expression vs variability

```{r}
cur_mus <- mus[,paste("S", 1:14, sep = "")]
cur_mus <- cur_mus[rownames(cur_eps),]

ggplot(data.frame(mus = log(rowMeans(cur_mus)),
                  eps = rowMeans(cur_eps, na.rm = TRUE))) + 
  geom_point(aes(mus, eps))

highVar.lowMean <- rownames(cur_mus)[log(rowMeans(cur_mus)) < -1 &
                                         rowMeans(cur_eps, na.rm = TRUE) > 0.5]

highVar.highMean <- rownames(cur_mus)[log(rowMeans(cur_mus)) > 1 &
                                         rowMeans(cur_eps, na.rm = TRUE) > 0.5]

lowVar.highMean <- rownames(cur_mus)[log(rowMeans(cur_mus)) > 1 &
                                         rowMeans(cur_eps, na.rm = TRUE) < 0]

lowVar.lowMean <- rownames(cur_mus)[log(rowMeans(cur_mus)) < -1 &
                                         rowMeans(cur_eps, na.rm = TRUE) < 0]
```

## Visualie Peaks for these categories

```{r}
cur_genes <- highVar.lowMean
cur_tss <- tss[tss@elementMetadata$gene_id %in% cur_genes]

cur_mat = normalizeToMatrix(cur_file, cur_tss, 
                        value_column = "score",
                        extend = 5000, mean_mode = "w0", 
                        w = 50, keep = c(0,0.99))

median(rowMeans(cur_mat))

EnrichedHeatmap(cur_mat, col = c("white", "red", "black"))

cur_genes <- highVar.highMean
cur_tss <- tss[tss@elementMetadata$gene_id %in% cur_genes]

cur_mat = normalizeToMatrix(cur_file, cur_tss, 
                        value_column = "score",
                        extend = 5000, mean_mode = "w0", 
                        w = 50, keep = c(0,0.99))

median(rowMeans(cur_mat))

EnrichedHeatmap(cur_mat, col = c("white", "red", "black"))

cur_genes <- lowVar.lowMean
cur_tss <- tss[tss@elementMetadata$gene_id %in% cur_genes]

cur_mat = normalizeToMatrix(cur_file, cur_tss, 
                        value_column = "score",
                        extend = 5000, mean_mode = "w0", 
                        w = 50, keep = c(0,0.99))

median(rowMeans(cur_mat))

EnrichedHeatmap(cur_mat, col = c("white", "red", "black"))

cur_genes <- lowVar.highMean
cur_tss <- tss[tss@elementMetadata$gene_id %in% cur_genes]

cur_mat = normalizeToMatrix(cur_file, cur_tss, 
                        value_column = "score",
                        extend = 5000, mean_mode = "w0", 
                        w = 50, keep = c(0,0.99))

median(rowMeans(cur_mat))

EnrichedHeatmap(cur_mat, col = c("white", "red", "black"))
```

# Clustering of variability profiles

```{r}
# Cluster genes based on variability profiles
cur_eps <- cur_eps[!is.na(rowMeans(cur_eps)),]
clusters <- kmeans(cur_eps, centers = 20)

# Form average profiles of clusters
mat <- matrix(data = NA, nrow = 20, ncol = ncol(cur_eps))
colnames(mat) <- colnames(cur_eps)

for(i in 1:20){
  mat[i,] <- colMeans(cur_eps[clusters$cluster == i,])
  plot(mat[i,])
}

# Visualize peaks
for(i in 1:20){
  cur_genes <- rownames(cur_eps)[clusters$cluster == i]
  cur_tss <- tss[tss@elementMetadata$gene_id %in% cur_genes]

  cur_mat = normalizeToMatrix(cur_file, cur_tss, 
                        value_column = "score",
                        extend = 5000, mean_mode = "w0", 
                        w = 50, keep = c(0,0.99))

  print(EnrichedHeatmap(cur_mat, col = c("white", "red", "black")))
}


```