---
title: "Variability over pseudotime"
author: "Nils Eling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, 
        encoding = encoding, output_file = '/Users/nils/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/Variability/VariabilityOverPT.html') })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I ran the BASiCS model on each cell population individually using the replicates to estimate technical variance. For each gene I fitted a linear regression between each iteration against the group number as response variable. This gives us a post-hoc posterior distribution of the slope of the linear model.

# Variability over pseudotime

```{r}
# Load libraries and data
library(BASiCS)
library(coda)
library(pheatmap)
library(openxlsx)
library(cowplot)
library(ggplot2)
library(rtracklayer)
library(EnrichedHeatmap)
library(GenomicFeatures)
library(EnsDb.Mmusculus.v79)
library(scater)
library(plyr)
library(reshape2)
library(pheatmap)
library(viridis)
source("../../Functions/auxiliary.R")

# Read in sce
sce <- readRDS("/Users/nils/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/data/10X_data/SCE_all_clusters.rds")
sce <- sce[,grepl("B6", colData(sce)$Sample) &
             colData(sce)$AnnotatedClusters %in% 
             c("S1", "S2", "S3", "S4", "S5", "S6", "S7", "S8", "S9")]
sce <- normalize(sce)

# Generate feature annotation
prom <- promoters(genes(EnsDb.Mmusculus.v79),
                  upstream = 1500, downstream = 500)
prom <- prom[seqnames(prom) %in% c(as.character(1:19), "X", "Y", "MT"),]
tss <- promoters(genes(EnsDb.Mmusculus.v79), 
                 upstream = 0, downstream = 1)
tss <- tss[seqnames(tss) %in% c(as.character(1:19), "X", "Y", "MT"),]
seqlevels(tss) <- c(as.character(1:19), "X", "Y", "MT")

# Read in blacklist file
blacklist <- import("/Users/nils/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/Cnr/Blacklist_mm10.bed")
seqlevels(blacklist) <- sub("^chr", '', seqlevels(blacklist))

# Read in gene names
genenames <- read.table("/Users/nils/Google Drive File Stream/My Drive/Christina/Tc1_single_cell/data/raw_reads/10X/do15983/Mus_human_chr21/genes.tsv", sep = "\t", stringsAsFactors = FALSE)
rownames(genenames) <- genenames[,1]

# Read in post-hoc distribution of slopes based on epsilons
fits.epsilon <- readRDS("/Users/nils/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/10X/Variability/data/fits_epsilon_B6.rds")
fits.mu <- readRDS("/Users/nils/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/10X/Variability/data/fits_mu_B6.rds")

# Remove coloumns containins NAs
select <- !is.na(colSums(fits.epsilon)) & !is.na(colSums(fits.mu))
fits.epsilon <- fits.epsilon[,select]
fits.mu <- fits.mu[,select]

# Visualize some trace plots for the MCMCs
plot(mcmc(fits.epsilon[,1]))
plot(mcmc(fits.epsilon[,10]))
plot(mcmc(fits.epsilon[,100]))
plot(mcmc(fits.epsilon[,1000]))
```

The trace plots show convergence for the slope - we can use this for posterior tail propability testing.

# Testing for fits on epsilons

```{r}
# Estimate the EFDR
TestSlope.eps <- BASiCS:::HiddenThresholdSearchTestDE(ChainLFC = fits.epsilon, 
                                        Epsilon = 0,
                                        ProbThreshold = NULL, GenesSelect = NULL,
                                        EFDR = 0.1, Task = "Slope testing")

OptThreshold.eps <- TestSlope.eps$OptThreshold[1]

slope.eps <- colMedians(fits.epsilon)
```

# Testing for fits on mus

```{r}
# Estimate the EFDR
TestSlope.mu <- BASiCS:::HiddenThresholdSearchTestDE(ChainLFC = fits.mu, 
                                        Epsilon = 0,
                                        ProbThreshold = 0.8, GenesSelect = NULL,
                                        EFDR = 0.1, Task = "Slope testing")

OptThreshold.mus <- TestSlope.mu$OptThreshold[1]

slope.mu <- colMedians(fits.mu)

# Build output dataframe
df.out <- data.frame(row.names = colnames(fits.mu),
                     Genenames = genenames[colnames(fits.mu),2],
                     Slope.mu = slope.mu,
                     Probability.mu = TestSlope.mu$Prob,
                     Result.mu = ifelse(TestSlope.mu$Prob > OptThreshold.mus & 
                                       slope.mu > 0, "MeanUp",
                                     ifelse(TestSlope.mu$Prob > OptThreshold.mus & 
                                              slope.mu < 0, "MeanDown", "NoDiff")),
                     Slope.epsilon = slope.eps,
                     Probability.epsilon = TestSlope.eps$Prob,
                     Result.epsilon = ifelse(TestSlope.eps$Prob > OptThreshold.eps & 
                                       slope.eps > 0, "VariabilityUp",
                                     ifelse(TestSlope.eps$Prob > OptThreshold.eps & 
                                              slope.eps < 0, "VariabilityDown", "NoDiff")))

# Order genes based on their increase in variability over PT
df.out <- df.out[order(df.out$Slope.epsilon, decreasing = TRUE),]
df.out$regulation <- paste(df.out$Result.mu, 
                           df.out$Result.epsilon, sep = "_and_")

write.xlsx(df.out, "/Users/nils/Dropbox (Cambridge University)/SST_spermatocytes/Analysis/10X/Results/VariabilityOverPT.xlsx")
```

# Classification of genes into 8 groups

- Increase in variability, no change in mean expression
- Decrease in variability, no change in mean expression
- Increase in variability, increase in mean expression
- Decrease in variability, increase in mean expression
- Increase in variability, decrease in mean expression
- Decrease in variability, decrease in mean expression
- No change in variability, increase in mean expression
- No change in variability, decrease in mean expression
- No change in variability, no change in mean expression


```{r}
ggplot(df.out) + geom_point(aes(Slope.mu, Slope.epsilon, colour = regulation)) +
  scale_color_manual(values = c("#fddbc7", "#d7301f", "#7f0000", "#d1e5f0", "#0570b0", "#081d58", "#f7fcfd", "#f4a582", "#92c5de")) + 
  xlab("Slope, Mean expression vs PT") + ylab("Slope, Variability vs PT")
```

# Order cells along the principle curve

```{r}
# Compute HVG
HVgenes <- HVG(sce)

# Compute PCA
pca <- prcomp(t(logcounts(sce)[HVgenes,]))

# Pseudo rank
prank <- PT(rd = pca$x[,1:2], clusters = colData(sce)$AnnotatedClusters,
            col_vector = metadata(sce)$color_vector)

# DPT 
#dpt <- diffusionPT(sce, HVgenes, clusters = colData(sce)$AnnotatedClusters,
#            col_vector = metadata(sce)$color_vector)
```

# Visualize example genes per category

```{r}
# Increase in mean expression, no change in variability
head(df.out[order(df.out$Slope.mu, decreasing = TRUE),], 50)

# Gpx4
cur.gene <- "Gpx4"
cur.Z <- (logcounts(sce)[rowData(sce)$Symbol == cur.gene,] -
             mean(logcounts(sce)[rowData(sce)$Symbol == cur.gene,]))/
  sd(logcounts(sce)[rowData(sce)$Symbol == cur.gene,])

Up_NoDiff <- ggplot(data.frame(PT = prank[,"lambda"],
                  gene = cur.Z,
                  cell_type = colData(sce)$AnnotatedClusters)) + 
  geom_point(aes(PT, gene, colour = cell_type)) + 
  xlab("Pseudotime") + ylab(paste(cur.gene, ", z-score", sep = "")) +
  geom_abline(slope = 0, intercept = 0, colour = "dark red") + 
  scale_color_manual(values = metadata(sce)$color_vector) + 
  guides(color = FALSE)

```
